---
title: "DATA 609 HW 3"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ex. 1
Write down Newton's formula for finding the minimum of $f(x) = \frac{3x^4 - 4x^3}{12}$
Solution:
Finding the minimum is equivalent to finding the roots for f'(x)=0. 
Newton's Method can solve this numerically using the iterative step of:
$$x_{k+1}=x_k - \frac{f'(x_k)}{f''(x_k)} $$
In our case we have
$$x_{k+1}=x_k - \frac{x_k^2(x_k -1)}{3x_k^2-2x_k)}  $$
Implementing this algorithm in R
```{r}
dfdx <- function(x) {
    x^2 * (x - 1)}

df2dx <- function(x) {
    3 * x^2 - 2 * x}

init <- 3
tol <- .00001

newton <- function(dfdx, df2dx, init, tol) {
    x = init
    while (abs(dfdx(x)) > tol) {
        x = x - dfdx(x)/df2dx(x)
        }
    x
    }

newton(dfdx, df2dx, init, tol)
```


## Ex. 2

Explore optimize() in R and try to solve the previous problem.
```{r}
f <- function (x) (3*x^4 - 4*x^3)/12
xmin <- optimize(f, c(-10, 10), tol = .00001)
xmin$minimum
```


## Ex. 3
Use any optimization algorithm to find the minimum of $f(x,y)= (x-1)^2 +100(y-x^2)^2$ in the domain
$-10 \le x,y \le 10$. Discuss any issues concerning the optimization process.
```{r}
library(GA)
fxy <- function(x, y)
{
  (x - 1)^2 + 100*(y - x^2)^2
}

x <- y <- seq(-10, 10, by = 0.1)
f <- outer(x, y, fxy)
persp3D(x, y, f, col.palette = bl2gr.colors)

GA <- ga(type = "real-valued", 
         fitness =  function(x) -fxy(x[1], x[2]),
         lower = c(-10, -10), upper = c(10, 10), 
         popSize = 100, maxiter = 1000, run = 100, seed = 160)
summary(GA)
plot(GA)
```
I did not really encounter any issues using the GA package to optimize this problem. I did find it hard to get the solution closer than 3 decimal places even after increasing the maximum iterations and the run parameter (which exits the algorithm after no solution improvement after that many iterations)

## Ex. 4
Explore the optimr package for R and try to solve the previous problem.

```{r}
library(optimr)
fxy2 <- function(par)
{
  (par[1] - 1)^2 + 100*(par[2] - par[1]^2)^2
}

optimr(par=c(.5,.5), fxy2, lower=c(-10,-10), upper=c(10,10), method='L-BFGS-B')
```
It is interesting to note how much faster this bounded BFGS method converged to the solution than the genetic algorithm.