---
title: "Data 609 - HW 2"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ex. 1 
Show $x^2 +e^x +2x^4 +1$ is convex.

Solution:

Since the non negative weighted sums of convex functions are also convex (from 2.1.3) 
I must show that the 3 separate functions of $f_1=x^2$, $f_2 = e^x$ and $f_3 = 2x^4 +1$ are convex.
I then have $\alpha f_1 + \beta f_2$ with $\alpha,\beta = 1 \ge 0$. 
This definition of non negative weighted sums can then be recursively applied to a
combination of any number of convex functions which in this case is 3.

From 2.1.2 we have the definition of a convex function where a function is convex if:

$f(\alpha x+ \beta y \le \alpha f(x) + \beta f(y)$    $\forall x,y$ and s.t. 

$\alpha \ge 0, \beta \ge 0, \alpha + \beta = 1$

Show $f_1=x^2$ is convex:

$(\alpha x + \beta y)^2 \le \alpha x^2 + \beta y^2 \implies$ expanding and moving all variables to one sides

$\alpha^2x^2 + 2\alpha\beta xy+\beta^2 y^2 -\alpha x^2 - \beta y^2 \le 0$ now factoring terms gives

$\alpha x^2(\alpha-1) + 2\alpha \beta xy +\beta y^2(\beta-1)\le 0$ substituting in for 1 which by definition is equal to $\alpha + \beta \implies$

$\alpha \beta x^2 +2\alpha\beta xy + \alpha \beta y^2\le 0$ factoring yields

$\alpha \beta(2xy-x^2-y2) \le 0 \implies -\alpha\beta(x-y)^2 \le0 \implies$

$\alpha\beta(x-y)^2 \ge 0$ which must be true because by definition $\alpha,\beta \ge 0$ and $(x-y)^2 \ge 0$   $\forall x,y \in R$

Show $f_2 = e^x$ is convex: See figure 2.1 on p.21

Show $f_3 = 2x^4 +1$ is convex:

$f_3$ is convex because its the composition of a affine function of the form $Ax +b$ with 
$g = x^4$ which is strictly increasing.


## Ex. 2
Show that the mean of the exponential distribution $p(x) = \lambda e^{-\lambda x} , x \ge 0 (\lambda > 0)$
is $\mu = 1\lambda$ and variance is $\sigma ^2 = 1\lambda ^2$

The mean and variance are given by the moment generating function. For the exponential distribution the moment is given by:

$M(t) = \frac{\lambda}{\lambda - t}$ 

Therefore, $M'(t)=\frac{\lambda}{(\lambda - t)^2}$ 
 and the mean $\mu = M'(0)=\frac{1}{\lambda}$. 
 
 The variance is given by $M''(0)-M'(0)^2$
 Since, $M''(t)= \frac{2\lambda}{(\lambda - t)^3}$ the variance is $\frac{1}{\lambda^2}$

## Ex. 3
It is estimated that there is a typo in every 250 data entries in a database, assuming the number of typos can obey the Poisson distribution. For a given 1000 data entries,

The probability of exactly 4 typos:
```{r}
dpois(x=4, lambda=4)
```

The probability of no typo at all:
```{r}
dpois(x=0, lambda=4)
```

Draw 1000 samples with $\lambda=4$ and show their histogram.
```{r}
Poisson <- rpois(1000, lambda=4)
hist(Poisson)
```


