---
title: "Data 609 - HW 4"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r}
library("ggplot2") 
```


## Ex. 1
For example 19 on page 79 in the book, carry out the regression using R.


```{r}
x <- c(-.98, 1, 2.02, 3.03, 4)
y <- c(2.44, -1.51, -.47, 2.54, 7.52)
lm1 <- lm(y ~ x)
summary(lm1)
```
So we have a simple linear model with $\beta_0 = .4038$ and $\beta_1 = .9373$ having x account for .263 of the variability of y.

## Ex. 2

Implement the nonlinear curve-fitting of example 20 on page 83 for the following data:

```{r}
x <- c(.1, .5, 1, 1.5, 2, 2.5)
y <- c(.1, .28, .4, .4, .37, .32)
df <- data.frame(x, y)
fx <- function(x, a, b){
  x / (a + (b * x^2))}

nlm1 <- nls(y ~ fx(x, a, b), data = df , start = list(a = 1, b = 1))
summary(nlm1)
```
So the non-linear curve that fits these data points is:
$$y = \frac{x}{1.48544 + 1.00212x^2}$$

We can verify by plotting the fitted function and the original data points.

```{r}

data_fun <- data.frame(x = - 10:10,          
                       y = c(fx(- 10:10, 1.48544, 1.00212)))

ggplot(data_fun, aes(x, y)) +
  geom_line() +
  geom_point(data = df, aes(x = x, y = y), color = "blue")
  
```
The points are reasonably close to the fitted non linear function.


## Ex. 3

For the data with binary y values, try to fit the following data to the nonlinear function
$$ y = \frac{1}{1 + e^{a +bx}}$$
starting with a = 1 and b = 1.

```{r}
x <- c(.1, .5, 1, 1.5, 2, 2.5)
y <- c(0, 0, 1, 1, 1, 0)
df2 <- data.frame(x, y)
fx2 <- function(x, a, b){
  1 / (1 + exp(a + b * x))}
nlm2 <- nls(y ~ fx2(x, a, b), data = df2, start = list(a = 1, b = 1))
summary(nlm2)
```
```{r}

data_fun <- data.frame(x = - 10:10,          
                       values = c(fx2(- 10:10, 36.42, -48.51)))

ggplot(data_fun, aes(x, values)) +
  geom_line() +
  geom_point(data = df2, aes(x = x, y = y), color = "blue")
  
```

It does not appear that the binary variables given can possibly be fit to the function $y = \frac{1}{1 + e^{a +bx}}$